|

### 2021-02-13 figure.Rmd
Create figure.Rmd
edits: Rmd output directory

---
title: "Figures"
output:
  pdf_document: default
---

@@

```{r, fig.height=6, fig.width=8, fig.cap="(A) Framework for generating theoretical branching river networks. Branching river networks are depicted as a network of connected habitat patches, in which the number of habitat patches $N_p$ and branching probability $P_{b}$ dictate the ecosystem size and complexity ($N_p = 50$ and $P_{b} = \\{0.2, 0.5, 0.8\\}$ in this example). Environmental conditions at headwaters (i.e., the most upstream patches) are drawn randomly from a normal distribution and propagate downstream with local environmental noise (a spatially autoregressive process with white noise; Methods). Habitat patches are colored in proportion to environmental values (similar colors have similar environmental values). (B) Example of intensively surveyed watersheds in Hokkaido, Japan (the red-colored watershed in C). Red dots indicate sampling sites for fish surveys. (C) Map of study regions (left, Hokkaido, Japan; right, Midwest, US). Watersheds (i.e., metacommunities) are gray-shaded in proportion to the number of sampling sites."}

>>>>

---
title: "Figures"
output:
  pdf_document: default
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

@@

```{r, fig.height=6, fig.width=8, fig.cap="(A) Framework for generating theoretical branching river networks. Branching river networks are depicted as a network of connected habitat patches, in which the number of habitat patches $N_p$ and branching probability $P_{b}$ dictate the ecosystem size and complexity ($N_p = 50$ and $P_{b} = \\{0.2, 0.5, 0.8\\}$ in this example). Environmental conditions at headwaters (i.e., the most upstream patches) are drawn randomly from a normal distribution and propagate downstream with local environmental noise (a spatially autoregressive process with white noise; Methods). Habitat patches are colored in proportion to environmental values (similar colors have similar environmental values). In this example, we used values of environmental variation at headwaters $\\sigma_{h} = 1$ and the degree of local environmental noise $\\sigma_{l} = 0.01$. (B) Example of intensively surveyed watersheds in Hokkaido, Japan (the red-colored watershed in C). Red dots indicate sampling sites for fish surveys. (C) Map of study regions (left, Hokkaido, Japan; right, Midwest, US). Watersheds (i.e., metacommunities) are gray-shaded in proportion to the number of sampling sites."}

### 2021-02-13 manuscript.Rmd
edits: Rmd output directory

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
output:
  html_notebook: default
  word_document: default
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous ecological pattern[@lomolinoEcologyMostGeneral2000] and has formed the basis of area-based conservation strategies[@desmetUsingSpeciesareaRelationship2004; @neigelSpeciesareaRelationshipsMarine2003]. However, many ecosystems possess a complex spatial structure that cannot be represented by area (the scale-invariant complexity)[@rodriguez-iturbeFractalRiverBasins2001; @turnerLandscapeEcologyTheory2015], and its role in scaling biodiversity is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbellUnifiedNeutralTheory2001; @kallimanisHowDoesHabitat2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of spatial difference in species composition was apparent when simulated species dispersal is limited. Cross-watershed comparisons across two geographic regions (the Hokkaido island in Japan and the midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.
bibliography: reference.bib
csl: nature.csl
---

@@

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity[@turnerLandscapeEcologyTheory2015;@rodriguez-iturbeFractalRiverBasins2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodriguez-iturbeFractalRiverBasins2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodriguez-iturbeFractalRiverBasins2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grantLivingBranchesPopulation2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure[@carraraDendriticConnectivityControls2012; @faganConnectivityFragmentationExtinction2002; @teruiMetapopulationStabilityBranching2018;@brownDendriticNetworkStructure2010], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@rinaldoRiverNetworksEcological2018;@petersonModellingDendriticEcological2013]. For example, this research has highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches[@carraraDendriticConnectivityControls2012]. However, most research has explored the consequences of branching complexity for local community structure[@brownDendriticNetworkStructure2010; @carraraDendriticConnectivityControls2012] or has relied solely on theoretical arguments with limited replications of branching architecture[@holtRoleBranchingMaintenance2018]. At present, we lack a comprehensive evaluation of how branching complexity scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services[@dudgeonFreshwaterBiodiversityImportance2006a].

Here, we hypothesize that ecosystem size and complexity dictate biodiversity patterns in rivers, a prime example of complex branching ecosystems. The recurrent merging of diverse streams and rivers produces larger, ecologically distinct channels, ultimately forming a fluvial network with fractal branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should control the ecosystem's habitat heterogeneity (habitat diversity per unit area). Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity). We predict that ecosystem size and branching complexity increase watershed-scale species richness ($\gamma$ diversity) by enhancing local species richness ($\alpha$ diversity) and/or spatial difference of species composition ($\beta$ diversity) under different ecological scenarios. The present study combines theory and empirical analysis of extensive community data from two different regions of the globe to provide crucial insights into how ecological communities are structured in complex branching networks.

@@

We depicted branching ecosystems as a bifurcating network of connected habitat patches[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$[@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution)[@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns[@peckhamReformulationHortonLaws1999]. An alternative option is the use of optimal channel networks[@carraroGenerationApplicationRiver2020]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to link theoretical predictions with empirical observations.

@@

where $r_{ix}(t)$ is the reproductive rate of species $i$ given the environmental condition at patch $x$ and time $t$, $r_{0,i}$ the maximum reproductive rate of species $i$, $K_{x}$ the carrying capacity at patch $x$, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, and $S$ the number of species in a metacommunity. $K_x$ was expressed as a function of the number of upstream habitat patches $A_x$ ($K_x = 100A_x$) to represent increasing habitat size with increasing upstream watershed area[@altermattRiverNetworkProperties2013; @teruiThreeEcologicalFactors2016]. The parameter $\alpha_{ij}$ is expressed as the strength of interspecific competition relative to that of intraspecific competition such that interspecific competition is greater than intraspecific competition if $\alpha_{ij}$ > 1 (intraspecific competition coefficient was set constant as $\alpha_{ii} = 1$). $\alpha_{ij}$ was drawn randomly from a uniform distribution as $\alpha_{ij} \sim Unif(0, \alpha_{max})$. The density-independent reproductive rate $r_{ix}(t)$ is affected by abiotic environments (non-consumable) and determined by the following Gaussian function:

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). These watersheds are small enough (`r round(min(c(wsd_subset[[1]]$area, wsd_subset[[2]]$area)))` to `r round(max(c(wsd_subset[[1]]$area, wsd_subset[[2]]$area)))` km^2^) to assume that fishes can disperse therein at an ecological time scale[@comteFishDispersalFlowing2018; @giamEnvironmentPredationGovern2016a], while posing challenges to traverse across watersheds (the ocean or lentic habitats). Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating diversity metrics (**Supplementary Information**).

We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity[@chaoRarefactionExtrapolationHill2014; @chaseEcosystemDecayExacerbates2020]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT'[@hsiehINEXTPackageRarefaction2016]. We assessed the completeness of the observed species richness based on sample coverage, which is the proportion of observed species richness to asymptotic species richness. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity[@whittakerVegetationSiskiyouMountains1960].

>>>>

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
output:
  word_document: default
  html_notebook: default
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous pattern[@lomolinoEcologyMostGeneral2000] and has formed the basis of area-based conservation strategies[@desmetUsingSpeciesareaRelationship2004; @neigelSpeciesareaRelationshipsMarine2003]. However, many ecosystems possess a complex spatial structure that cannot be represented by area (the scale-invariant complexity)[@rodriguez-iturbeFractalRiverBasins2001; @turnerLandscapeEcologyTheory2015], and its role in scaling biodiversity is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbellUnifiedNeutralTheory2001; @kallimanisHowDoesHabitat2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of spatial difference in species composition was apparent when simulated species dispersal is limited. Cross-watershed comparisons across two geographic regions (the Hokkaido island in Japan and the midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.
bibliography: reference.bib
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

@@

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity[@turnerLandscapeEcologyTheory2015;@rodriguez-iturbeFractalRiverBasins2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodriguez-iturbeFractalRiverBasins2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodriguez-iturbeFractalRiverBasins2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grantLivingBranchesPopulation2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure[@carraraDendriticConnectivityControls2012; @faganConnectivityFragmentationExtinction2002; @teruiMetapopulationStabilityBranching2018;@brownDendriticNetworkStructure2010], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@rinaldoRiverNetworksEcological2018;@petersonModellingDendriticEcological2013]. For example, these studies have highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches[@carraraDendriticConnectivityControls2012]. However, most research has explored the consequences of branching complexity for local community structure[@brownDendriticNetworkStructure2010; @carraraDendriticConnectivityControls2012] or has relied solely on theoretical arguments with limited replications of branching architecture[@holtRoleBranchingMaintenance2018]. At present, we lack a comprehensive evaluation of how branching complexity scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services[@dudgeonFreshwaterBiodiversityImportance2006].

Here, we hypothesize that ecosystem size and complexity dictate biodiversity patterns in rivers - a prime example of complex branching ecosystems. The recurrent merging of diverse streams and rivers produces larger, ecologically distinct channels, ultimately forming a fluvial network with fractal branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should control the ecosystem's habitat heterogeneity (habitat diversity per unit area). Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity). We predict that ecosystem size and branching complexity increase watershed-scale species richness ($\gamma$ diversity) by enhancing local species richness ($\alpha$ diversity) and/or spatial difference of species composition ($\beta$ diversity) under different ecological scenarios. The present study combines theory and empirical analysis of extensive community data from two different regions of the globe to provide crucial insights into how ecological communities are structured in complex branching networks.

@@

We depicted branching ecosystems as a bifurcating network of connected habitat patches[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$[@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution)[@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns[@peckhamReformulationHortonLaws1999]. An alternative approach is the use of optimal channel networks[@carraroGenerationApplicationRiver2020]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to compare theoretical predictions with empirical observations.

@@

where $r_{ix}(t)$ is the reproductive rate of species $i$ given the environmental condition at patch $x$ and time $t$, $r_{0,i}$ the maximum reproductive rate of species $i$, $K_{x}$ the carrying capacity at patch $x$, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, and $S$ the number of species in a metacommunity. $K_x$ was expressed as a function of the number of upstream habitat patches $A_x$ ($K_x = 100A_x$) to represent increasing habitat size with increasing upstream watershed area[@altermattRiverNetworkProperties2013; @teruiThreeEcologicalFactors2016]. The parameter $\alpha_{ij}$ represents the strength of interspecific competition relative to that of intraspecific competition: interspecific competition is greater than intraspecific competition if $\alpha_{ij}$ > 1 (intraspecific competition coefficient was set constant as $\alpha_{ii} = 1$). $\alpha_{ij}$ was drawn randomly from a uniform distribution as $\alpha_{ij} \sim Unif(0, \alpha_{max})$. The density-independent reproductive rate $r_{ix}(t)$ is affected by abiotic environments (non-consumable) and determined by the following Gaussian function:

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). These watersheds are small enough (`r round(min(c(wsd_subset[[1]]$area, wsd_subset[[2]]$area)))` to `r round(max(c(wsd_subset[[1]]$area, wsd_subset[[2]]$area)))` km^2^) to assume that fishes can disperse therein at an ecological time scale[@comteFishDispersalFlowing2018; @giamEnvironmentPredationGovern2016], while posing challenges to traverse across watersheds (the ocean or lentic habitats). Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating diversity metrics (**Supplementary Information**).

We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity[@chaoRarefactionExtrapolationHill2014a; @chaseEcosystemDecayExacerbates2020]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT'[@hsiehINEXTPackageRarefaction2016]. We assessed the completeness of the observed species richness based on sample coverage, which is the proportion of observed species richness to asymptotic species richness. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity[@whittakerVegetationSiskiyouMountains1960].

### 2021-02-13 supplementary_information.Rmd
edits: Rmd output directory

---
title: |
  | Supplementary Information for:
  | Ecosystem size and complexity dictate riverine biodiversity
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
output:
  pdf_document:
    toc: yes
    toc_depth: 2
bibliography: reference.bib
csl: nature.csl
---

>>>>

---
title: |
  | Supplementary Information for:
  | Ecosystem size and complexity dictate riverine biodiversity
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
output:
  pdf_document:
    toc: yes
    toc_depth: 2
bibliography: reference.bib
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---
